Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics

### 2024-11-08 GroupPolarizationStatmod/paper/ms.tex
updates to Discussion

Ordinal scales are an indispensable tool for psychological research on
how social influence changes people's beliefs, opinions, knowledge, or
similar psychological content. We show here that we must be careful not
to mistake the measurement procedure as psychological reality when
designing experiments.

Replication crisis, generalizability crisis \autocite{Yarkoni2022}, and
general lack of coordination in asking the right questions to
efficiently understand the mechanisms behind . Understanding mechanisms
is the goal of science \autocite{Machamer2000,Craver2006}. However, in
social and psychological sciences, finding mechanisms is more difficult
due to the complexity of social psychological systems
\autocite{Dubova2023}, which require a high density of conceptual
definitions that may be difficult to harmonize.

>>>>

\mt{Restate results and importance and outline this section.}

\mt{One more brick to fall in the ongoing replication and generalizability crisis \autocite{Yarkoni2022}.}

Since there seems to be no selection pressure on false discovery rates, 
experiment design for studying group polarization was free to evolve randomly.
Various choices for the number of bins and other instrumental details may have
then become entrenched in different communities defined by 
common theoretical perspectives taken by researchers in those communities. 
Therefore, when 
Isenberg (1986)~\cite{Isenberg1986} reports that his meta-analysis found more
support for the XX theory than the YY theory in the form of greater effect sizes,
it may just be random cultural drift that the XX community instruments were
more prone to greater .

This leads to one of many follow-up research questions generated by this work.
First, are there general principles of experimental design that could reduce
the false detection rate for group polarization or similar experimental protocols?
Second, what is the best way to calculate false detection rate? 
Likely, even if we cannot identify parameters consistent with simple consensus
that generate spurious group polarization shifts in the large $N$ case, one
may still guess at some simple consensus parameters and caluclate the false
detection rate as we did. On the flip side, we likely found only one of infinitely
many simple consensus parameters that could result in a false detection. Does this
have an effect on the calculated false positive rate? More generally, our
work motivates similar analyses of sub-fields of social 
psychology and that do not account for Likert-based experimental designs in
their statistical analyses. We provided a template for this future work.

\mt{concluding paragraph: first sentences summarize the Discussion, then
in 1-2 sentences summarize the paper.}
